---
title: "Madrid train bombings"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
author: "Mohamed Ali"
date: "10/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggraph)
library(igraph)
library(visNetwork)
library(seriation)
library(plotly)
library(plotly)
library(dplyr)
library(tourr)
link<-read.table("trainData.dat", quote="\"", comment.char="")
node<-read.table("trainMeta.dat", quote="\"", comment.char="")
```

# Network visualization of terrorist connections



```{r 1.1, echo=FALSE}
node$id<-rownames(node)
node<-node[,c(3,1,2)]
colnames(node)[1:3]<-c("id","name","part")
node$part[node$part==1]<-"Participated in placing the explosives"
node$part[node$part==0]<-"Did't participated"
colnames(link)[1:3]<-c("from","to","weight")
link <- link[order(link$from, link$to),]
node$group<-node$part
node$label<-node$name
net <- graph_from_data_frame(d=link, vertices=node, directed=T)
node$value<-strength(net)
link$width=link$weight


visNetwork(node,link)%>%visLegend()%>%visOptions(highlightNearest = list(enabled = T, degree = 1, hover = T))%>%visPhysics(solver = "repulsion")
```

### What we can see ?
The graph above shows a network of network of the individuals suspected of having participated in the bombing of commuter trains in Madrid on March 11, 2004, data clustered by bombing group and connection between nodes made base on the strength of tie between individuals, analyzing the graph manually we can observe 4 groups, the first one (the most obvious one) which is at the center of the graph, which consist of individuals participated directly in the bombing incident “colored in blue” and their connection with individuals who did not participate “colored in yellow” with a different tie of strength,  the second cluster at the top of the graph which include only individuals who didn’t participate but the have a multiple connection ties. the third cluster is not obvious but to some extend we can confirm on it which is the cluster the left bottom of the graph, which consist of yellow group (not-participated) individuals with a limited connection with blue group (participated individuals). Finally, the fourth group which are the outliers the suspected individuals with no connection ties.

### Individual has the best opportunity to spread the information in the network

```{r 1.2, echo=FALSE}
visNetwork(node,link)%>%visLegend()%>%visOptions(highlightNearest = list(enabled = T, degree = 2, hover = T))%>%visPhysics(solver = "repulsion")

```

### hummm!!
The graph above highlights nodes with connection path length 1 or 2, data shows that individuals with high number of connections have better opportunity to spread the information in the network, graph shows that Jamal Zougam as the defendant with most connection ties with other individuals.
Jamal Zougam Moroccan-Spanish 48-year-old convicted of being a perpetrator of the March 11, 2004, train bombings in Madrid, Spain, that killed 191 people and injured more than 1,800. According to counter extremism website Zougam who had previously been brought to the attention of authorities for his links to al-Qaeda members, placed at least one of the bombs.  He also owned the shop where the mobile phones used to detonate the bombs were sold (https://www.counterextremism.com/extremists/jamal-zougam). The authorities found that Zougam have a strong connection with Imad Eddin Barakat Yarkas (alias Abu Dahdah), a Spaniard prosecuted for his involvement in the 9/11 attacks, which can be seen obviously in the network graph in the connection tie.
The graph shows another suspect connection ties are similar to Zougam.  Mohamed Chaoui who also accused in the attempt of Madrid. According to the Guardian Mr Chaoui's name features on a phone tap of an alleged Spanish al-Qaida cell that may have helped prepare the September 11 plot in addition to that he also charged with multiple counts of murder, attempted murder, stealing a vehicle, belonging to a terrorist organization and four counts of carrying out terrorist acts. The BBC reported on 19th of March that Mohamed Bekkali also been areasted, our graph shows also who a strong connection tie between him and both Zougam and Chaoui.http://news.bbc.co.uk/2/hi/europe/3597885.stm
Naima Oulad Akcha, a non-directly participated individual in Madrid bombing also arrested according to BBC as the only woman suspect, however her connection ties shows that she maybe on of the main participants on the attacks.
Interestingly, four of the suscepts who were accused of placing the bomb (Abddenabi Koujam, Mohamed Oulad Akcha, Rachid Oulad Akcha, and Nasredine Boushoa) have a very small connection ties compared with other individuals, reports online did not shows much information about their roles in the operation.


## More investigation clusters by optimizing edge betweenness
```{r 1.3, echo=FALSE,warning=FALSE}
node1<-node
net <- graph_from_data_frame(d=link, vertices=node, directed=F)
ceb <- cluster_edge_betweenness(net) 
node1$group<-ceb$membership
visNetwork(node1,link)%>%visIgraphLayout()%>%visOptions(highlightNearest = list(enabled = T, degree = 2, hover = T))
```

### Let's see 
Above graph represent clusters obtained by optimizing edge betweenness, the results show 5 main clusters, the main basic clusters centered in the middle of the graphs has been mentioned in Q1, however there was a non-defined cluster from step 1 which is the in between cluster at the top left of the graph. Outliers have been defined a one cluster in step 1, while in on this method each outlier been defined and single clusters.

## Maybe a heatmap!! performing permutation by Hierarchical Clustering
```{r 1.4, echo=FALSE}
netm <- get.adjacency(net, attr="width", sparse=F)
colnames(netm) <- V(net)$name
rownames(netm) <- V(net)$name
rowdist<-dist(netm)
order1<-seriate(rowdist, "HC")
ord1<-get_order(order1)
reordmatr<-netm[ord1,ord1]
plot_ly(z=~reordmatr, x=~colnames(reordmatr), 
        y=~rownames(reordmatr), type="heatmap")
```

### A heatmap !! why not, let's see
Above graph represent heatmap obtained by adjacency matrix representation to perform permutations and Hierarchical Clustering (HC) seriation method, the graph highlights obvious cluster at the top right, this cluster is similar to some extend to clusters obtained in step 1 & 3, however it’s not clear the connection ties between other suspects expect of those with high strength of connection ties for example, Jamal Zougam and Imad Eddin Barakat Yarkas.
